# Chapter 4 - Statistical Inference

```{r}
# Set up libraries and functions
library(tidyverse)
library(rosdata)
library(rstanarm)
library(patchwork)
options(mc.cores = parallel::detectCores())
```

```{r}
n = 1000
y = 700

estimate <- y / n
se <- sqrt(estimate*(1-estimate)/n)
int_95 <- estimate + qnorm(c(0.025, 0.975))*se
```

```{r}
# SE of difference
x <- rnorm(1000000, mean = 0, sd = 4)
y <- rnorm(1000000, mean = 0, sd = 5)

sd(x)
sd(y)

sd(x - y)

sqrt(sd(x)^2 + sd(y)^2)
```

```{r}
y <- c(35, 34, 38, 35, 37)
n <- length(y)
estimate <- mean(y)
se <- sd(y) / sqrt(n)
int_50 <- estimate + qt(c(0.25, 0.75), n-1)*se
int_95 <- estimate + qt(c(0.025, 0.975), n-1)*se
```

```{r}
y <- rep(c(0,1,2,3,4), c(600,300,50,30,20))
n <- length(y)
estimate <- mean(y)
se <- sd(y) / sqrt(n)
int_50 <- estimate + qt(c(0.25, 0.75), n-1)*se
int_95 <- estimate + qt(c(0.025, 0.975), n-1)*se
```

```{r}
int_95 <- estimate + qt(c(0.025, 0.975), n-1)*se
int_95*1000000
```

# Exercises

## 1.)

```{r}
incentive = 500
no_incentive = 500
incentive_response_rate = 0.5
no_incentive_response_rate = 0.4

# Average treatment effect
ate = incentive_response_rate - no_incentive_response_rate

# SE for incentive
incentive_SE <- sqrt((incentive_response_rate*(1-incentive_response_rate) / incentive))

# SE for no incentive
no_incentive_SE <- sqrt((no_incentive_response_rate*(1-no_incentive_response_rate) / no_incentive))

# SE of difference
sqrt(incentive_SE^2 + no_incentive_SE^2)
```

The SE of the ATE of 10 percent is ~ 3.1%

## 2.)

Assuming that men and women are equal in the population:

The standard error for each proportion is: $SE = \sqrt{((p(1-p)) / (n)})$

The quantity we want is the difference between support for the candidate for men and women,

and the formula for the difference is $SE = \sqrt{SE_{men}^2} + SE_{women}^2})$

So for the SE to be less than 5, let's consider the value that will maximise the variance (0.5 for both).

then 0.0025 >= 2*SE^2_men

So we want the SE for each sex to be less than 0.03535534

Now we can work out the sample size for this:

0.03535534 = sqrt(0.25 / n)
0.03535534^2 = 0.25 / n
n = 0.25 / 0.03535534^2

```{r}
n = 0.25 / 0.03535534^2
```

assuming equal proportions in the population, we have n = 200 for men and n = 200 for women, so we want to sample at least 400 people.

## 3.)

*Shooter A*:
n_A = 20
prop_A = 0.3

*Shooter B*:
n_B = 20
prop_B = 0.4

```{r}
# Approach 1 - simulation
shots_A <- rbinom(100000, 20, 0.3)
shots_B <- rbinom(100000, 20, 0.4)

mean(shots_B > shots_A)
```

The probability that the better shooter makes more shots is about 70 percent.

## 4.)

Let's say that we want only a 5% chance of being wrong, or in other words, a 95% chance of being correct.

Then we want the probability that shots_B > shots_A at least 95%.

```{r}
# Create function to compute probability that B shoots more than A based on simulation
prob_b_greater_a <-
    function(n) {
        shots_A <- rbinom(100000, n, 0.3)
        shots_B <- rbinom(100000, n, 0.4)
        
        mean(shots_B > shots_A)
    }

# Increase n from 20 until response reaches 0.95
n_trials = 20
while(prob_b_greater_a(n_trials) < 0.95) {
    n_trials = n_trials + 1
    print(n_trials)
}

paste("So we have to perform", n_trials + 1, "trials to properly distinguish the two shooters")
```

## 5.)

```{r}
data("iris")

iris <-
    iris |> 
    arrange(Sepal.Length)

sample_means <- c()

for (i in 1:500) {
sample_mean <-
    iris |> 
slice_sample(n = 20) |> 
    pull(Sepal.Length) |> 
    mean()
sample_means[i] <- sample_mean
}

hist(sample_means)
```

## 6.)

```{r}
girls <-
    c(
        0.4777,
        0.4875,
        0.4859,
        0.4754,
        0.4874,
        0.4864,
        0.4813,
        0.4787,
        0.4895,
        0.4797,
        0.4876,
        0.4859,
        0.4857,
        0.4907,
        0.5010,
        0.4903,
        0.4860,
        0.4911,
        0.4871,
        0.4725,
        0.4822,
        0.4870,
        0.4823,
        0.4973
    )

observed_SD <- sd(girls)

# Compare this to what would be expected from the binomial distribution
expected_SD <- sqrt(0.4857*(1-0.4857) / 3900)

expected_SD - observed_SD

# Here we see that the observed SD is slightly smaller than what would be expected.

# Now lets compare our observed variance with what is expected under the chi squared distribution with 23 degrees of freedom
observed_var <- var(girls)

sampleMean <- mean(girls)
theoreticalVar <- expected_SD^2
Q <- 23 * observed_var / theoreticalVar
pValue <- pchisq(Q, df=23, lower.tail = TRUE)
    
```

So there is insufficient evidence to reject the randomness model under the standard hypothesis testing framework.

